Live probes for free

### 2023-07-19 paper/example-programming.tex
fix perf

%Shape: Rectangle [id:dp48530454252997635] 
\draw   (11,13.3) -- (68.6,13.3) -- (68.6,34.5) -- (11,34.5) -- cycle ;
%Shape: Rectangle [id:dp4741172814816692] 
\draw   (11,34.5) -- (68.6,34.5) -- (68.6,89.7) -- (11,89.7) -- cycle ;
%Shape: Rectangle [id:dp39333640528563685] 
\draw   (83.8,13.3) -- (141.4,13.3) -- (141.4,34.5) -- (83.8,34.5) -- cycle ;
%Shape: Rectangle [id:dp852338699133017] 
\draw   (83.8,34.5) -- (141.4,34.5) -- (141.4,89.7) -- (83.8,89.7) -- cycle ;
%Shape: Rectangle [id:dp7978026394194542] 
\draw   (155.6,13.3) -- (213.2,13.3) -- (213.2,34.5) -- (155.6,34.5) -- cycle ;
%Shape: Rectangle [id:dp7423976709217543] 
\draw   (155.6,34.5) -- (213.2,34.5) -- (213.2,89.7) -- (155.6,89.7) -- cycle ;
%Shape: Rectangle [id:dp7923849318907303] 
\draw   (225.8,13.7) -- (283.4,13.7) -- (283.4,34.9) -- (225.8,34.9) -- cycle ;
%Shape: Rectangle [id:dp8325747263531937] 
\draw   (225.8,34.9) -- (283.4,34.9) -- (283.4,90.1) -- (225.8,90.1) -- cycle ;
%Shape: Rectangle [id:dp3066063334990279] 
\draw   (343,14.1) -- (400.6,14.1) -- (400.6,35.3) -- (343,35.3) -- cycle ;
%Shape: Rectangle [id:dp9425547748588214] 
\draw   (343,35.3) -- (400.6,35.3) -- (400.6,90.5) -- (343,90.5) -- cycle ;
%Straight Lines [id:da5537198175803346] 
\draw    (69.4,56.1) -- (80,56.1) ;
\draw [shift={(83,56.1)}, rotate = 180] [fill={rgb, 255:red, 0; green, 0; blue, 0 }  ][line width=0.08]  [draw opacity=0] (7.14,-3.43) -- (0,0) -- (7.14,3.43) -- cycle    ;
%Straight Lines [id:da9992180079914922] 
\draw    (141.4,56.1) -- (152,56.1) ;
\draw [shift={(155,56.1)}, rotate = 180] [fill={rgb, 255:red, 0; green, 0; blue, 0 }  ][line width=0.08]  [draw opacity=0] (7.14,-3.43) -- (0,0) -- (7.14,3.43) -- cycle    ;
%Straight Lines [id:da24526356738515775] 
\draw    (212.4,56.1) -- (223,56.1) ;
\draw [shift={(226,56.1)}, rotate = 180] [fill={rgb, 255:red, 0; green, 0; blue, 0 }  ][line width=0.08]  [draw opacity=0] (7.14,-3.43) -- (0,0) -- (7.14,3.43) -- cycle    ;
%Straight Lines [id:da7346663349503276] 
\draw [line width=1.5]    (284.4,56.1) -- (337.6,56.47) ;
\draw [shift={(341.6,56.5)}, rotate = 180.4] [fill={rgb, 255:red, 0; green, 0; blue, 0 }  ][line width=0.08]  [draw opacity=0] (9.29,-4.46) -- (0,0) -- (9.29,4.46) -- cycle    ;

% Text Node
\draw (15.2,17) node [anchor=north west][inner sep=0.75pt]   [align=left] {line : 2};
% Text Node
\draw (22.4,39) node [anchor=north west][inner sep=0.75pt]   [align=left] {n : 3};
% Text Node
\draw (88,17) node [anchor=north west][inner sep=0.75pt]   [align=left] {line : 3};
% Text Node
\draw (95.2,39) node [anchor=north west][inner sep=0.75pt]   [align=left] {n : 3};
% Text Node
\draw (94.8,53.2) node [anchor=north west][inner sep=0.75pt]   [align=left] {i \ : 0};
% Text Node
\draw (159.8,17) node [anchor=north west][inner sep=0.75pt]   [align=left] {line : 4};
% Text Node
\draw (167,39) node [anchor=north west][inner sep=0.75pt]   [align=left] {n : 3};
% Text Node
\draw (166.6,53.2) node [anchor=north west][inner sep=0.75pt]   [align=left] {i \ : 0};
% Text Node
\draw (230,17) node [anchor=north west][inner sep=0.75pt]   [align=left] {line : 3};
% Text Node
\draw (237.2,39) node [anchor=north west][inner sep=0.75pt]   [align=left] {n : 3};
% Text Node
\draw (236.8,53.6) node [anchor=north west][inner sep=0.75pt]   [align=left] {i \ : 1};
% Text Node
\draw (347.2,17) node [anchor=north west][inner sep=0.75pt]   [align=left] {line : 6};
% Text Node
\draw (354.4,39) node [anchor=north west][inner sep=0.75pt]   [align=left] {n : 3};
% Text Node
\draw (354,54) node [anchor=north west][inner sep=0.75pt]   [align=left] {i \ : 2};
    \end{tikzpicture}
  \end{minipage}
  \caption{Stack Recording Example}
  \label{fig:stack-recording}
\end{figure}

@@

\begin{itemize}
  \item What is the average performance during a live programming session?
  \item What is the compilation time as a function of programme size ?
  \item How long does it take to load the code into the debugger, depending on the size of the programme?
  \item How long does it need to run the program, depending on the number of steps to make the stack recording ?
\end{itemize}

@@

To measure the performance of our approach in the general context of a live programming session, we measured 4 points :

%In the 3 languages, we have implemented time measurement for the binary_search method (see appendix) with the array \[1,2,3,4,5,6\] looking for 9.
%The execution was repeated 100 times for each language.

@@

\subsubsection{What is the compilation time as a function of programme size ?}
\subsubsection{How long does it take to load the code into the debugger, depending on the size of the programme?}

@@

>>>>

%Shape: Rectangle [id:dp48530454252997635] 
\draw   (11,13.3) -- (68.6,13.3) -- (68.6,34.5) -- (11,34.5) -- cycle ;
%Shape: Rectangle [id:dp4741172814816692] 
\draw   (11,34.5) -- (68.6,34.5) -- (68.6,89.7) -- (11,89.7) -- cycle ;
%Shape: Rectangle [id:dp39333640528563685] 
\draw   (83.8,13.3) -- (141.4,13.3) -- (141.4,34.5) -- (83.8,34.5) -- cycle ;
%Shape: Rectangle [id:dp852338699133017] 
\draw   (83.8,34.5) -- (141.4,34.5) -- (141.4,89.7) -- (83.8,89.7) -- cycle ;
%Shape: Rectangle [id:dp7978026394194542] 
\draw   (155.6,13.3) -- (213.2,13.3) -- (213.2,34.5) -- (155.6,34.5) -- cycle ;
%Shape: Rectangle [id:dp7423976709217543] 
\draw   (155.6,34.5) -- (213.2,34.5) -- (213.2,89.7) -- (155.6,89.7) -- cycle ;
%Shape: Rectangle [id:dp7923849318907303] 
\draw   (225.8,13.7) -- (283.4,13.7) -- (283.4,34.9) -- (225.8,34.9) -- cycle ;
%Shape: Rectangle [id:dp8325747263531937] 
\draw   (225.8,34.9) -- (283.4,34.9) -- (283.4,90.1) -- (225.8,90.1) -- cycle ;
%Shape: Rectangle [id:dp3066063334990279] 
\draw   (343,14.1) -- (400.6,14.1) -- (400.6,35.3) -- (343,35.3) -- cycle ;
%Shape: Rectangle [id:dp9425547748588214] 
\draw   (343,35.3) -- (400.6,35.3) -- (400.6,90.5) -- (343,90.5) -- cycle ;
%Straight Lines [id:da5537198175803346] 
\draw    (69.4,56.1) -- (80,56.1) ;
\draw [shift={(83,56.1)}, rotate = 180] [fill={rgb, 255:red, 0; green, 0; blue, 0 }  ][line width=0.08]  [draw opacity=0] (7.14,-3.43) -- (0,0) -- (7.14,3.43) -- cycle    ;
%Straight Lines [id:da9992180079914922] 
\draw    (141.4,56.1) -- (152,56.1) ;
\draw [shift={(155,56.1)}, rotate = 180] [fill={rgb, 255:red, 0; green, 0; blue, 0 }  ][line width=0.08]  [draw opacity=0] (7.14,-3.43) -- (0,0) -- (7.14,3.43) -- cycle    ;
%Straight Lines [id:da24526356738515775] 
\draw    (212.4,56.1) -- (223,56.1) ;
\draw [shift={(226,56.1)}, rotate = 180] [fill={rgb, 255:red, 0; green, 0; blue, 0 }  ][line width=0.08]  [draw opacity=0] (7.14,-3.43) -- (0,0) -- (7.14,3.43) -- cycle    ;
%Straight Lines [id:da7346663349503276] 
\draw [line width=1.25, dashed]    (284.4,56.1) -- (337.6,56.47) ;
\draw [shift={(341.6,56.5)}, rotate = 180.4] [fill={rgb, 255:red, 0; green, 0; blue, 0 }  ][line width=0.08]  [draw opacity=0] (9.29,-4.46) -- (0,0) -- (9.29,4.46) -- cycle    ;

% Text Node
\draw (15.2,17) node [anchor=north west][inner sep=0.75pt]   [align=left] {line : 3};
% Text Node
\draw (22.4,39) node [anchor=north west][inner sep=0.75pt]   [align=left] {n : 3};
% Text Node
\draw (88,17) node [anchor=north west][inner sep=0.75pt]   [align=left] {line : 4};
% Text Node
\draw (95.2,39) node [anchor=north west][inner sep=0.75pt]   [align=left] {n : 3};
% Text Node
\draw (94.8,53.2) node [anchor=north west][inner sep=0.75pt]   [align=left] {i \ : 0};
% Text Node
\draw (159.8,17) node [anchor=north west][inner sep=0.75pt]   [align=left] {line : 5};
% Text Node
\draw (167,39) node [anchor=north west][inner sep=0.75pt]   [align=left] {n : 3};
% Text Node
\draw (166.6,53.2) node [anchor=north west][inner sep=0.75pt]   [align=left] {i \ : 0};
% Text Node
\draw (230,17) node [anchor=north west][inner sep=0.75pt]   [align=left] {line : 4};
% Text Node
\draw (237.2,39) node [anchor=north west][inner sep=0.75pt]   [align=left] {n : 3};
% Text Node
\draw (236.8,53.6) node [anchor=north west][inner sep=0.75pt]   [align=left] {i \ : 1};
% Text Node
\draw (347.2,17) node [anchor=north west][inner sep=0.75pt]   [align=left] {line : 7};
% Text Node
\draw (354.4,39) node [anchor=north west][inner sep=0.75pt]   [align=left] {n : 3};
% Text Node
\draw (354,54) node [anchor=north west][inner sep=0.75pt]   [align=left] {i \ : 2};
    \end{tikzpicture}
  \end{minipage}
  \caption{Stack Recording Example}
  \label{fig:stack-recording}
\end{figure}

@@

\begin{itemize}
  \item What is the average performance during a live programming session?
  \item How long does it take to load the code into the debugger, depending on the size of the programme?
  \item How long does it take to run the program, depending on the number of stack frames needed to make the stack recording?
\end{itemize}

@@

@@

To evaluate the performance of our approach in the general context of a live programming session, we measured the times taken by the different stages of the interaction loop:

For the 3 languages we have implemented, we have performed the measurements for the execution of a binary search function\ref*{fig:binary-search}.
For each language, the agent was initialized, then the code was compiled, loaded into the debugger and executed (with the array [1,2,3,4,5,6] and target 9) 100 times.
Each execution generate a stack recording of approximately 20 stack frames.
The results are shown in table \ref*{table:average-time}.

\subsubsection{How long does it take to load the code into the debugger, depending on the size of the programme?}

@@

\newpage
\section*{Appendix}
\label{sec:appendix}

\begin{figure}[h]
  \centering
    \begin{lstlisting}[language=C]
int binary_search(int arr[], int length, int target) {
  int left = 0;
  int right = length - 1;
  while (left <= right) {

      int mid = (left + right) / 2;
      if (arr[mid] == target) {
          return mid;
      } else if (arr[mid] < target) {
          left = mid + 1;
      } else {
          right = mid - 1;
      }
  }
  return 0;
}
    \end{lstlisting}

    \begin{lstlisting}[language=Python]
def binary_search(arr, target):
    left = 0
    right = len(arr) - 1

    while left <= right:
        mid = (left + right) // 2

        if arr[mid] == target:
            return mid
        elif arr[mid] < target:
            left = mid + 1
        else:
            right = mid - 1

    return 0 
    \end{lstlisting}

    \begin{lstlisting}[language=Java]
public class BinarySearch {
    public static int binarySearch(int[] array, int key){
        int low = 0;
        int high = array.length - 1;
        while (low <= high) {
            int mid = (low + high) / 2;
            int value = array[mid];
            if (value < key) {
                low = mid + 1;
            } else if (value > key) {
                high = mid - 1;
            } else {
                return mid;
            }
        }
        return 0;
    }
}
      \end{lstlisting}
  \caption{Binary Search in C, Python and Java}
  \label{fig:binary-search}
\end{figure}
